---
title: "Data Science in a Day"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(dplyr)
```


# Overview

## Aim
Teach a bit of everything along the data science process.

## Methodology
Rapid iterations of:

- lecture
- structured exercises
- apply learnings to our project

## Resources
- Use R from your browser (see business card)
- Get the materials at [bit.ly/dsiadrepo](http://bit.ly/dsiadrepo)
- Get the slides at [bit.ly/dsiadslides](http://bit.ly/dsiadslides)

## Agenda
- Identifying the business challenge
- Converting a business challenge into a data science problem
- Building a plan of attack
- Preparing data
- Preparing predictive columns
- Building candidate models
- Evaluating models
- Selecting models against business criteria
- Putting a model into production

## Abraham Lincoln
> Give me six hours to chop down a tree and I will spend the first four sharpening the axe.

## 6Ps
> Proper Planning Prevents Piss-Poor Performance

# The business challenge

## Why have a challenge to solve?
Data science is expensive, shouldn't do it without good reason

## What sort of challenges might a business have?
Most business challenges align to one or more of the three parts of boosting profits:
- By increasing revenue 
- By retaining revenue sources
- By reducing overheads

## What business challenge to start with?
- Pick **low effort** and **medium/high** value
- Don't pick something business critical
- One that's measurable - if it ain't measured it ain't science!

## Formulating the business challenge
Need a succinct description that contains key measures of success e.g.

> We would like increase the amount our customers spend with us and our level of integration with the customers' lives by increasing the number of our products that  customers consume.

## Exercise
Formulate a business challenge for your organisation.

## Chicago City Council

Chicago tourism has taken a recent hit because of a spate of food poisonings at Chicago food establishments. What can we do to address this issue?

# Business challenge to data science problem
## Identify behaviours
What behaviour(s) by customers, staff, third parties contribute to the business goal?

## Manipulating behaviours
In what ways could these behaviours be influenced to impact the key metrics?

## Match the right prediction
What predictions or prescriptions could you make that would be the influence needed?

## Example
> To increase customers consumption of multiple products, customers need to see relevant products with a speedy purchase option. We need to identify for each customer what relevant is, where possible, to ensure custommers are most likely to purchase.

## Exercise
Addressing your business problem, what could you influence with data science?

## Chicago City Council

# Building a plan of attack
## How will you know you've won?
First and foremost you make sure you can measure your impact.

## What sort of actions will need to be taken?
What will be the operational constraints your work will need to work within? 

It's important to know if you have to have something that works on devices, can be done by batch, or requires regulatory approval.

## What sort of data could be involved?
What data sets might you need for exploratory analysis and later modelling?

It's a good idea to think about the retrieval and processing required of these in light of your operational constraints.

## Project planning
How much time can you dedicate to a prototype and/or building a full model? Who will put things into production?

Get some of these things straight to help you avoid going into an analysis rabbit hole.

## Ethics
Would your work disadvantage a group of people based on things like gender, race, disability? If yes, should you proceed?

## Exercise
Work through our points of consideration for your candidate data science project. What sort of constraints would be in place? What data do you have available? Could you measure your impact?

## Chicago City Council

# Preparing data
## Data usage rights
Are you allowed to use data for the purposes you're proposing to put it to?

## Get data
We need some data for our project. 

We usually need enough data to take a sample to develop our model on and another set to test our model. I'd suggest sample sizes of no less than a thousand.

If you want to be able to predict missing values, the amount of data required increases.

## Notebook time!
[studio.azureml.net](https://studio.azureml.net/)

# Data I/O

# Data manipulation basics
```{r child="subDocs/dplyrbasics.Rmd"}
```
