

## Filtering - Conditions
```{r}
iris %>%
  filter(Species == "setosa")
```

## Filtering - Positions
```{r}
iris %>%
  slice(which.max(Sepal.Length)) 
```

## Column manipulation - Selection
```{r}
iris %>%
  select(Sepal.Length, Sepal.Width, Species)
```

## Column manipulation - Selection
```{r}
iris %>%
  select(-Sepal.Length, -Sepal.Width)
```

## Column manipulation - Selection
```{r}
iris %>%
  select(starts_with("S"))
```

## Column manipulation - Aliasing

```{r}
iris %>%
  rename(sepal_length = Sepal.Length, sepal_width = Sepal.Width) %>%
  head
```


## New columns

```{r}
iris %>%
  mutate(Sepal.Area= Sepal.Length*Sepal.Width)
```

## New columns

```{r}
iris %>%
  mutate(Sepal.Area1 = Sepal.Length*Sepal.Width,
         Rounded.Sepal.Area1=round(Sepal.Area1))
```

## Aggregations
```{r}
iris %>%
  summarise(mean_sepal_length=mean(Sepal.Length),
            n=n())
```

## Groupings
```{r}
iris %>%
  group_by(Species )%>%
  summarise(mean_sepal_length=mean(Sepal.Length),
            n=n())
```

## Sorting
```{r}
iris %>%
  arrange(desc(Sepal.Length))
```

## Sampling
```{r}
iris %>% 
  sample_n(2)
```

## Joins

```{r}
irisLookup<-data.frame( Species = c("Other","setosa","virginica") ,
                        Property = c(1,1,2)
                      )
```

## Left join
```{r, warning=FALSE}
iris %>%
  left_join(irisLookup) %>%
  group_by(Species) %>%
  summarise(N=n(),
            RHS_NAs = sum(is.na(Property)))
```

## Inner join
```{r, warning=FALSE}
iris %>%
  inner_join(irisLookup) %>%
  group_by(Species) %>%
  summarise(N=n(),
            RHS_NAs = sum(is.na(Property)))
```

## Full join
```{r, warning=FALSE}
iris %>%
  full_join(irisLookup) %>%
  group_by(Species) %>%
  summarise(N=n(),
            RHS_NAs = sum(is.na(Property)))
```

## Anti-join
```{r, warning=FALSE}
iris %>%
  anti_join(irisLookup) %>%
  group_by(Species) %>%
  summarise(N=n())
```

## Semi-join
```{r, warning=FALSE}
iris %>%
  semi_join(irisLookup) %>%
  group_by(Species) %>%
  summarise(N=n())
```

## Joins on different names
```{r, warning=FALSE}
irisLookup<-data.frame( species = c("Other","setosa","virginica") ,
                        Property = c(1,1,2)
                      )

iris %>%
  inner_join(irisLookup, by= c("Species"="species")) %>%
  group_by(Species) %>%
  summarise(N=n(),
            RHS_NAs = sum(is.na(Property)))
```

## Joins on compound keys
```{r, warning=FALSE}
irisLookup<-data.frame( Species = c("Other","setosa","virginica") ,
                        Sepal.Length = c(5.9, 4.9, 4.7)
                      )

iris %>%
  inner_join(irisLookup, by= c("Species","Sepal.Length")) %>%
  group_by(Species) %>%
  summarise(N=n())
```

## Joins on calculated values
```{r, warning=FALSE}
irisLookup<-data.frame( Species = c("Other","setosa","virginica") ,
                        Sepal.Length = c("Large","Large","Small")
                      )

iris %>%
  mutate(Sepal.Bracket = ifelse(Sepal.Length>mean(Sepal.Length), "Large", "Small")) %>%
  inner_join(irisLookup, by= c("Species","Sepal.Bracket"="Sepal.Length")) %>%
  group_by(Species) %>%
  summarise(N=n())
```
